### 2023-02-13 article.tex
Use Great Britain English

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analyzed the reproducibility of registered reports
in the field of psychology, 
where 60\% of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For articles in Science magazine, 12\% of 204
studies published the code \cite{stodden2018empirical} 
(note that these were years 2011-2012).
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{1189px-Eukaryote_DNA-en.png}
  \caption{
    A cell has a nucleus that contains chromosomes. 
    Each of these chromosomes (46 in humans) consist out of DNA. 
    DNA itself consists out of 4 nucleotides, 
    as depicted by the horizontal sticks 
    with the colors red, yellow, green and blue.
    From \cite{sponk2012}
  }
  \label{fig:eukakyote_dna}
\end{figure}
\fi

@@

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a.png}
  \caption{
    Gene expression result due to an off-by-one error.
    The cells in the main table show how much mRNA is produced per cell 
    line (the columns) for different proteins (the rows).
    The colors above the cells, with the dendrogram, shows
    red for cells that do not respond (red) and do respond (blue) to an 
    antibiotic.
    The purple rectangles show samples that are duplicated.
    Note that, due to this, some
    samples that behave identically are both non-responders and responders.
    From \cite{baggerly2009deriving}
  }
  \label{fig:baggerly2009deriving}
\end{figure}
\fi

@@

\subsection{Containerization}

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
(however, that article was written before virtualization existed).
Virtualization allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster),
as are commonly needed in the field of genetic epidemiology experiments.

@@

For code to be ideal paradata, 
its creation and documentation need to align, 
\emph{not least because they shape each other} \cite{huvila2022improving}.
This emphasized part of the quote resonates strongly with the idea of 
literate programming \cite{knuth1984literate},
in which documentation and code is developed hand-in-hand.
Literate programming is the practice of writing a document mixed with code that 
will be run.
Contemporary examples of this idea are, among others, 
vignettes \cite{wickham2015r} (for the R programming language \cite{r}) 
and Jupyter notebooks \cite{wang2020assessing} 
(for the Julia \cite{Julia-2017}, Python \cite{van1995python} and R \cite{r}
programming languages).
rOpenSci, a community that, among others, reviews programming
code (\cite{ram2013ropensci,ram2018community}), is 
one example of where literate programming is enforced:
code must have extensive runnable examples
as part of the documentation \cite{ropensci_2021_6619350}.
Developing code and documentation at the same time 
leads to both better code and documentation \cite{reenskaug1989environment}.
In general, when developing software, 
it is recommended to 
to write documentation while writing software,
as well as to include many examples \cite{lee2018ten}.

@@

The percentage of (lines of) code tested is called the code coverage.
Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
The non-profit organization rOpenSci \cite{ram2018community},
which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.
CI opens up the possibility to informally annotate the
code coverage. 
For example, build badges
\iffalse
(e.g. as shown in Figure \ref{fig:badge_codecov})
\fi
are visual indications on a website that can, among others,
show a certain R package's code has a 100\% code coverage.

@@

The world of science would be a more open, humble, trustworthy, truthful
and helpful would the code that accompanies a scientific paper
be treated like a first class citizen. 
Doing so, however, is yet to be rewarded
and still both of the two scientists at the start of this paper 
can provide a good rationale for their behavior.
This will change when reward incentives are put into place 
that reward making paradata useful.
For code specifically, in any computational field,
the rewards are even higher, as reproducibility should again be 
a cornerstone in science.

@@

\begin{table}[h]
  \begin{tabular}{|p{2cm}|l|}
    \hline
    \textbf{Type of code} & \textbf{Recommendation} \\
    \hline
    No code       & Disallow papers that do not supply code publicly \\
    \hline
    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
    \hline
    Hosted code   & Use a standardized badge for build status \\
                  & Use a standardized badge for code coverage \\
                  & Standardize to link between paper and code host \\
    \hline
    Runnable code & Allow to upload containers without limits \\
                  & Website to host containers must allow for annotation \\
                  & Use a standardized badge for the container to be functioning  \\
                  & Standardize to link between container, paper and code repository \\
    \hline
  \end{tabular}
  \caption{Recommendations described in this paper}
  \label{tab:recommendations}
\end{table}

>>>>

To determine how useful code is, it needs to be available.
However, it is not common to publish the code of an experiment or analysis 
\cite{stodden2011trust,read2015sizing} (with a pleasant exception 
being \cite{conesa2019making}).
For example, in computer graphics, 
a field intimately familiar with computer code,
42\% of 454 SIGGRAPH papers supply computer code \cite{bonneel2020code}.
Another study analysed the reproducibility of registered reports
in the field of psychology, 
where 60\% of 62 studies supplied the code 
to redo the analysis \cite{obels2020analysis}.
For articles in Science magazine, 12\% of 204
studies published the code \cite{stodden2018empirical} 
(note that these were years 2011-2012).
Unpublished code has been the cause of some saddening examples,
such as an algorithm that detects breast cancer from images 
better than a human expert, 
yet failed to ever be reproduced \cite{haibe2020importance}.

@@

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{1189px-Eukaryote_DNA-en.png}
  \caption{
    A cell has a nucleus that contains chromosomes. 
    Each of these chromosomes (46 in humans) consist out of DNA. 
    DNA itself consists out of 4 nucleotides, 
    as depicted by the horizontal sticks 
    with the colours red, yellow, green and blue.
    From \cite{sponk2012}
  }
  \label{fig:eukakyote_dna}
\end{figure}
\fi

@@

\iffalse
\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\linewidth]{baggerly2009deriving_fig_1a.png}
  \caption{
    Gene expression result due to an off-by-one error.
    The cells in the main table show how much mRNA is produced per cell 
    line (the columns) for different proteins (the rows).
    The colours above the cells, with the dendrogram, shows
    red for cells that do not respond (red) and do respond (blue) to an 
    antibiotic.
    The purple rectangles show samples that are duplicated.
    Note that, due to this, some
    samples that behave identically are both non-responders and responders.
    From \cite{baggerly2009deriving}
  }
  \label{fig:baggerly2009deriving}
\end{figure}
\fi

@@

\subsection{Containerisation}

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.
This is close to the golden standard suggested by 
\cite{peng2011reproducible} 
\iffalse
(see also Figure \ref{fig:peng2011reproducible})
\fi
(however, that article was written before virtualisation existed).
Virtualisation allows one to provide a standard environment,
that can be used on any other computer.
The best known examples of container software are Docker and Singularity.
Due to access rights, Singularity is the only viable option to run
an experiment on a 
high performance computing (HPC) environment (e.g. a computer cluster),
as are commonly needed in the field of genetic epidemiology experiments.

@@

For code to be ideal paradata, 
its creation and documentation need to align, 
\emph{not least because they shape each other} \cite{huvila2022improving}.
This emphasised part of the quote resonates strongly with the idea of 
literate programming \cite{knuth1984literate},
in which documentation and code is developed hand-in-hand.
Literate programming is the practice of writing a document mixed with code that 
will be run.
Contemporary examples of this idea are, among others, 
vignettes \cite{wickham2015r} (for the R programming language \cite{r}) 
and Jupyter notebooks \cite{wang2020assessing} 
(for the Julia \cite{Julia-2017}, Python \cite{van1995python} and R \cite{r}
programming languages).
rOpenSci, a community that, among others, reviews programming
code (\cite{ram2013ropensci,ram2018community}), is 
one example of where literate programming is enforced:
code must have extensive runnable examples
as part of the documentation \cite{ropensci_2021_6619350}.
Developing code and documentation at the same time 
leads to both better code and documentation \cite{reenskaug1989environment}.
In general, when developing software, 
it is recommended to 
to write documentation while writing software,
as well as to include many examples \cite{lee2018ten}.

@@

The percentage of (lines of) code tested is called the code coverage.
Code coverage correlates with code quality \cite{horgan1994,del1995correlation}. 
The non-profit organisation rOpenSci \cite{ram2018community},
which peer-reviews code,
has made it a prerequisite to have a code coverage of 100\%.
CI opens up the possibility to informally annotate the
code coverage. 
For example, build badges
\iffalse
(e.g. as shown in Figure \ref{fig:badge_codecov})
\fi
are visual indications on a website that can, among others,
show a certain R package's code has a 100\% code coverage.

@@

The world of science would be a more open, humble, trustworthy, truthful
and helpful would the code that accompanies a scientific paper
be treated like a first class citizen. 
Doing so, however, is yet to be rewarded
and still both of the two scientists at the start of this paper 
can provide a good rationale for their behaviour.
This will change when reward incentives are put into place 
that reward making paradata useful.
For code specifically, in any computational field,
the rewards are even higher, as reproducibility should again be 
a cornerstone in science.

@@

\begin{table}[h]
  \begin{tabular}{|p{2cm}|l|}
    \hline
    \textbf{Type of code} & \textbf{Recommendation} \\
    \hline
    No code       & Disallow papers that do not supply code publicly \\
    \hline
    Raw code      & Don't publish raw code as text, host it on a code hosting website \\
    \hline
    Hosted code   & Use a standardised badge for build status \\
                  & Use a standardised badge for code coverage \\
                  & Standardise to link between paper and code host \\
    \hline
    Runnable code & Allow to upload containers without limits \\
                  & Website to host containers must allow for annotation \\
                  & Use a standardised badge for the container to be functioning  \\
                  & Standardise to link between container, paper and code repository \\
    \hline
  \end{tabular}
  \caption{Recommendations described in this paper}
  \label{tab:recommendations}
\end{table}
